\chapter{Contributors}

\section*{Audrey Hammer}

Audrey is finishing a BA in Philosophy at the University of Cambridge.
She is taking papers in Mathematical Logic, Philosophical Logic,
Wittgenstein and His Successors, and Kant. Her dissertation is on the
implications of independence and incompleteness in results on realism
set in theory. Next year she aims to pursue an MPhil in Philosophy at
the University of Cambridge, where she wants to explore the nature of
logic and mathematics.

\section*{Koda (Wenjing) Li}

Koda is a third-year Linguistics and Philosophy student at Brown
University. In philosophy, he is mainly interested in philosophy of
mind, philosophy of language, metaethics, and philosophical language.
In linguistics, he is mainly interested in semantics and its interface 

\section*{Wiktor Przybrorwski}

Wiktor is a third-year MA Economics and Philosophy student at the
University of St Andrews. His primary interests lie in the philosophy
of logic and language, particularly in topics such as naming and
reference, paradoxes, and non-classical logics.

\section*{Wilson Sugeng}

Wilson is an Economics and Philosophy undergraduate at the University of
St Andrews. He previously read Theology at Wheaton College, Illinois,
and spent a semester at Wycliffe Hall, Oxford. With work experience in
finance, microeconomic research, and church ministry, his research
interests include religious epistemology and the ways different academic
fields, religions, and philosophical traditions shape ethical and
metaphysical frameworks. Outside of academics, Wilson enjoys cooking and
hiking with the university's hillwalking club.
